---
title: "Feb 27 (Tue): Feature Engineering on Immigration Attitude"
output: html_notebook
---

# Step 1: import original data

```{r}
## 3:00 pm
library(foreign)
library(RStata)
library(tidyverse)
dataset <- read.dta("H:/DAT/BrExit/data/BES2017_W13_Panel_v1.2.dta")

```

# Step 2: what are the variables and how many of them
```{r}
demog_index <- grep("profile_",names(dataset))
pers_index <- grep("personality_",names(dataset))
vote_index <- grep("euRef",names(dataset))
immi_index <- grep("immig",names(dataset))

demog_feature <- names(dataset)[demog_index]
pers_feature <-names(dataset)[pers_index]
vote_features <-names(dataset)[vote_index]
immi_features <-names(dataset)[immi_index]
```

```{r}
dat_789 <- dataset %>%
  filter(wave7==1 & wave8 == 1 & wave9 ==1)

write.csv(dat_789, "dat_wave_789.csv")
```

```{r}
# condense dataset by removing the variables we are not looking at 
remove <- c("competent", "authLong", "grammar", "LookAfter", "overseasAid","^W1","^W2","^W3","^W4","^W5","^W6","^W10","^W11","^W12","^W13")
for(i in 1:length(remove)){
 index <-grep(remove[i], names(dataset))
 dat_789 <- dat_789 %>%
   select(-index)
}
```


Demographic:
profile_marital_statW11
profile_ethnicity (only after wave 10)
profile_work_statW12
gender
profile_household_children
housing
profile_newspaper
profile_gross_household
profile_gross_personal
ageGroup (10 yrs apart)
Age

"personality_agreeableness"     "personality_conscientiousness" "personality_extraversion"     
"personality_neuroticism"       "personality_openness" 


Political:
al_scale
lr_scale

Referendum:
postalTurnout
euRefVoteW
euSqueeze
euRefVotePostW8 (Which way did you vote? W8)

Immigration:
immigSelf
benefitsToMigrants
(immigType)
effectsEUImmigration

Other:
betterDealEU
ukCooperatePref
effectsEU2
effectsEU3
certaintyEUEffect