---
title: "BrExit Survey Data Exploration"
output:
  html_document: default
  html_notebook: default
---
## Week 3 Report
###Background:

British Election study: 
Examine the support of secessionist movements among first and secod generation immigrants. The data consists of answers to survey questions asked in different waves (total of 13), from 2014 - 2018. 

### Data source:


## Goals of previous week: 
###Success 
* Step 1. To visualize the general "opinion landscape" on leave/stay in the EU for each wave of survey. 
* Step 2. Further investigate the difference between sub demographic groups:  
    + Age_group
    + Education
    + Social Economic Status (?)
    + Urban rural
    + Party ID
    + Political philosophy  
    + Conservative -> break from EU
    + Pro-labour -> stay in EU
    
###Sweet 
* Create condensed data set to work with.
* Figure out accumulative weight.

###Amazing
* Investigate the general "opinion landscape" on leave/stay in the EU for intersection of waves.


```{r, message = FALSE, warning= FALSE, echo=FALSE}
# Read in
#library(readstata13)
library(tidyverse)
library(lubridate)
dat <- read.csv("wt_age_origin.csv")
```

### Data Background:
Original dataset has > 5000 variables, 68625 observations.
There are 13 waves of surveys, with ~3000 observations in each wave.

### Note:
Calculation of weights: \# of weights that said yes/ # of weights that said yes + said no 


## TODO: daily for wave 5 
## TODO: find wt_new_w10

## Preparing Data: create new dataset

The original dataset is person-level. It looks like: 
```{r, echo = FALSE}
head(dat[,c(2,50,37:49,3:5,71:75)])
```

```{r tidy = TRUE, echo = FALSE}
demo_dat <- dat %>% select(id,country,ageW1,countryOfBirth)
dat2 <- c() ## all responses
for(i in 1:13){
  if(i!= 5){
  wave_name <- paste0("^wave",i,"$")
  wave_col <- grep(wave_name, colnames(dat))
  
  vote_name <- paste0("^euRefVoteW",i,"$")
  vote_col<-grep(vote_name,colnames(dat))
  
  if(i<=9){
    wt_name <- paste0("^wt_core_W",i,"$")
    wt_col <- grep(wt_name,colnames(dat))
  }else if (i == 10){
    wt_name <- paste0("^wt_full_W",i,"$")
    wt_col <- grep(wt_name,colnames(dat))
  }else{
    wt_name <- paste0("^wt_new_W",i,"$")
    wt_col <- grep(wt_name,colnames(dat))
  }
  
  wave_i <- dat %>% 
    filter(dat[,wave_col]==1) %>%
    select(id, wt_col,vote_col, -wave_col) %>%
    mutate(wave = i) 
  
  names(wave_i) <- sub(wt_name, "wt", names(wave_i))
  names(wave_i) <- sub(vote_name, "vote", names(wave_i))
  
  dat2 <- rbind2(dat2,wave_i)
  }
}

ins_dat <- dat2 %>% 
  inner_join(demo_dat)
```

After manipulation, we get a wave-level data that looks like: 
```{r, echo = FALSE}
head(arrange(ins_dat, id)) 
```

#Step One: Plot general trend on timeline (with actual date/time)
```{r}
ins_dat <-ins_dat %>% mutate(endtime = ifelse(wave == 1, "2014-03-09",  ifelse (wave == 2, "2014-06-25",
                                    ifelse (wave == 3, "2014-10-17",
                                        ifelse (wave == 4, "2015-03-30",
                                            ifelse (wave == 5, "2015-05-06", 
                                               ifelse (wave == 6, "2015-05-26",
                                                  ifelse (wave == 7,"2016-05-4",
                                                     ifelse (wave == 8, "2016-06-22",
                                                       ifelse (wave == 9, "2016-07-04",
                                                         ifelse (wave == 10, "2016-12-12",
                                                             ifelse (wave == 11,"2017-05-03",
                                                                ifelse (wave == 12,"2017-06-07",
                                                                     "2017-06-23")))))))))))))

ins_dat$endtime <- as.Date(ins_dat$endtime)

ins_dat %>% group_by(endtime,vote)%>%
  summarise(count = sum(wt,na.rm = TRUE)) %>%
  ggplot(aes(x = endtime, y= count, group = vote, colour = vote))+geom_line() + geom_point() 
```


#Step Two: Look at demographic subgroups 

### Country of origin 

"For each demographic subgroup, how does the stay/leave opinion change over time?"
```{r tidy = TRUE, echo = FALSE, message=FALSE}
# number of responses for each subgroup in each wave

dat_origin <- ins_dat %>%
  ## re-categorize country of birth:
  mutate(countryOfBirth_General = ifelse(countryOfBirth %in% c("England","Northern Ireland","Republic of Ireland", "Wales"), "UK", countryOfBirth)) %>% 
  group_by(wave, countryOfBirth_General, vote) %>%
  summarise(group_vote = sum(wt, na.rm = TRUE)) %>%
  filter(vote == "Leave the EU" | vote == "Stay/remain in the EU") %>%
  spread(vote,group_vote) %>%
  plyr::rename(c("Stay/remain in the EU" = "Stay",
         "Leave the EU" = "Leave")) %>%
  mutate(stay_leave_ratio = Stay/Leave) %>%
  ggplot(aes(x = wave,y = stay_leave_ratio, group = countryOfBirth_General, colour = countryOfBirth_General)) + geom_point() + geom_line() +  ylab("Stay-leave ratio of the country of birth")+
  xlab("Across waves 1 ~ 13 (EXCEPT WAVE 5)") + 
  scale_x_continuous(breaks=seq(1,13))

# ## color
# color <- c("#b10026", "#fc4e2a", "#fd8d3c", "#969696",  "#252525","#238b45",
#            "#7570b3", "#e7298a", "#a6761d", "#d95f02")
# l2 <- c("Other: Commonwealth member country", 
#  "Other: European Union member country",
#  "Other: Rest of world",            
#  "Prefer not to answer",
#  NA,
# "Republic of Ireland",                
#  "Scotland",                        
#  "Wales",   
# "England",                           
# "Northern Ireland")
# 
# ## reordered the level
# dat_origin$countryOfBirth <- factor(dat_origin$countryOfBirth, levels = l2)


dat_origin %>%
  ggplot(aes(x = wave,y = stay_leave_ratio, group = countryOfBirth, colour = countryOfBirth)) + geom_point() + geom_line() +  ylab("Stay-leave ratio of the country of birth")+
  xlab("Across waves 1 ~ 13 (EXCEPT WAVE 5)") + 
  scale_x_continuous(breaks=seq(1,13)) #+   scale_colour_manual(values=color)


```

### Need further investigation: 
__Note: wierd drop in England origin__
```{r echo = FALSE, message = FALSE}
dat_leave <- ins_dat %>% 
  filter(vote == "Leave the EU") %>%
  select(-vote) %>%
  group_by(wave,countryOfBirth) %>%
  summarise(leave = sum(wt, na.rm = TRUE))
  
dat_stay <- ins_dat %>% 
  filter(vote == "Stay/remain in the EU") %>%
  select(-vote)%>%
  group_by(wave,countryOfBirth) %>%
  summarise(stay = sum(wt, na.rm = TRUE))

## Note: the % is over all 'Leave' voters in that wave 
dat_leave_total <- ins_dat %>%
  filter(vote == "Leave the EU") %>%
  group_by(wave) %>%
  summarise(leave_total = sum(wt, na.rm = TRUE))

dat_stay_total <- ins_dat %>%
  filter(vote == "Stay/remain in the EU") %>%
  group_by(wave) %>%
  summarise(stay_total = sum(wt, na.rm = TRUE))

dat_origin_p <- dat_leave_total %>%
  inner_join(dat_leave) %>%
  inner_join(dat_stay) %>%
  inner_join(dat_stay_total) %>%
  mutate(stay_p = stay/stay_total,
         leave_p = leave/leave_total)


 dat_origin_p %>%
  ggplot(aes(y = leave_p, x = wave, group = countryOfBirth, colour = countryOfBirth))+
  geom_line() + 
  geom_point() + 
  ylab("% of voters voted 'leave' of the country of birth")+
  xlab("Across waves 1 ~ 13 (EXCEPT WAVE 5)") + 
  scale_x_continuous(breaks=seq(1,13))+ theme(legend.position="none")+ coord_trans(y="log2")
```


```{r echo=FALSE}
 dat_origin_p %>%
  ggplot(aes(y = stay_p, x = wave, group = countryOfBirth, colour = countryOfBirth))+
  geom_line() + 
  geom_point() + 
  ylab("% of voters voted 'stay' of the country of birth")+
  xlab("Across waves 1 ~ 13 (EXCEPT WAVE 5)") + 
  scale_x_continuous(breaks=seq(1,13)) + coord_trans(y="log2")

#grid.arrange(leave_tot, stay_tot, nrow = 1)
```


### Examine opinion change of people with English origin 
```{r echo = FALSE}
#origin
En_vote <- ins_dat %>%
  filter(countryOfBirth == "England") %>%
  group_by(wave, vote) %>%
  summarise(wt_vote = sum(wt, na.rm = TRUE))

En_vote %>%
  ggplot(aes(x = wave, y = wt_vote, group = vote, color = vote)) +
  geom_point() + geom_line() + ggtitle("Country of Birth: 'England'")+ scale_x_continuous(breaks=seq(1,13)) + ylab("Votes (weighted raw counts)")
```

### Examine opinion change of people who don't want to reveal their country of origin
```{r echo = FALSE}
NA_vote <- ins_dat %>%
  filter(is.na(countryOfBirth)) %>%
  group_by(wave, vote) %>%
  summarise(wt_vote = sum(wt, na.rm = TRUE))

NA_vote %>%
  ggplot(aes(x = wave, y = wt_vote, group = vote, color = vote)) +
  geom_point() + geom_line() + scale_x_continuous(breaks=seq(1,13)) +ylab("Votes (weighted raw counts)") + ggtitle("Country of Birth: 'Prefer not to answer'")
```


## Examine opinion change of people who are from the rest of the world
```{r echo = FALSE}
In_vote <- ins_dat %>%
  filter(countryOfBirth == "Other: Rest of world") %>%
  group_by(wave, vote) %>%
  summarise(wt_vote = sum(wt, na.rm = TRUE))

In_vote %>%
  ggplot(aes(x = wave, y = wt_vote, group = vote, color = vote)) +
  geom_point() + geom_line() + ggtitle("Country of Birth: 'Rest of the world'") + scale_x_continuous(breaks=seq(1,13))
```

### Look into total demographic change (country of origin)
** Very wierd ** 
```{r  echo = FALSE}
 ins_dat %>%
  group_by(wave, countryOfBirth) %>%
  summarise(count = sum(wt, na.rm = TRUE)) %>%
  ggplot(aes(x = wave, y = count, group = countryOfBirth, color = countryOfBirth)) + geom_point() + geom_line() + ylab("Total response (weighted)") + ggtitle("Demographic change/check across survey waves") + coord_trans(y="log2")
```


### Next time: 
### Opinion divide within each group of country origin
```{r eval = FALSE, echo = FALSE}
dat_total <- ins_dat %>%
  group_by(wave, vote) %>%
  summarize(each_origin_total = sum(wt,na.rm=TRUE))%>%
  spread(vote,each_origin_total)

dat_ <- ins_dat %>% 
  group_by(wave,countryOfBirth,vote) %>%
  summarise(count = sum(wt, na.rm = TRUE))%>%
  inner_join(dat_total) %>%
  mutate(stay_p = stay/count,
         leave_p = leave/count) %>%
  gather(atti, percent, -countryOfBirth, -wave,-vote,)
```

```{r eval = FALSE, echo = FALSE}
  ggplot(aes(y = percent, x = wave,fill = atti))+ 
  geom_bar(stat="identity",position=position_dodge()) +
  ylab("% vote over total reponse (including 'I don't know')") +
  xlab("Across waves 1 ~ 13 (except 5)")
```


```{r eval = FALSE, echo = FALSE}

## Reduce to a condensed dataset (used code)
condensed <- dat[,c(1:67,72,75,index)]
write.csv(condensed, "wt_age_origin.csv")

index <-c()
for(i in 1:13){
  reg <- paste0("^euRefVoteW",i,"$")
  col_num<-grep(reg,colnames(dat))
  index <- c(index, col_num)
}

grep("^euRefVoteW1$", colnames(dat)) #265
```


## Next Goal  
* Mark 3 critical events 
* Continuous waves (1~13)
** Determine the most continuous waves (balance sample size)
* Other demographic groups

#TODO:  
  * missing vote for wave 5 ??? => check original dataset 
  * find wt_new_w10
  * Log scale for 
  * change in each wave

```{r eval = FALSE, echo = FALSE}
## create a dataset that is wave specific 
for(i in 1: 13){
  #wave_num <- paste0("wave",i);
  dat %>% filter(wave1 == 1, euRefVoteW1== "Stay/remain in the EU") %>% 
    summarise(stay = sum(wt_core_W1, na.rm = TRUE))
  
  dat %>% filter(wave1 == 1) %>% summarise(total = sum(wt_core_W1, na.rm = TRUE))
  
  dat %>% group_by()
}

t <- dat %>%
  filter(wave1==1) %>%
  group_by(euRefVoteW1) %>%
  summarise(opinion_count = sum(wt_core_W1, na.rm = TRUE))
  
  
t2 <- dat %>%  
  filter(wave1==1) %>%
  inner_join(t) %>%
  mutate(total_count = sum(wt_core_W1, na.rm = TRUE))%>%
  mutate(stay_p = opinion_count/total_count) %>%
  filter(euRefVoteW1 == "Stay/remain in the EU") %>%
  select(stay_p) %>%
  unique()

```



## Look at the sub-sample of people who are in all 13 waves
```{r}
all_13_id <- ins_dat %>% group_by(id) %>%
  summarise(num_wave = n())%>%
  filter(num_wave == 12) %>% ## SINCE wave 5 is ommited (So we have participants who just missed wave 5 too)
  select(id)

dat_13 <- ins_dat %>% ##65028 instances
  right_join(all_13_id)

# dat %>%
#   filter(wave1 == 1 & wave2 == 1 & wave3 == 1 & wave4 == 1 & wave5 == 1 & 
#            wave6 == 1 & wave7 == 1 & wave8 == 1 & wave9 == 1 & wave10 == 1&
#            wave11 == 1 & wave12 == 1 & wave13 == 1) %>%
#   group_by(wave, vote)


#test
ins_dat %>% 
        group_by(country,endtime,vote) %>%
        summarise(count = sum(wt, na.rm = TRUE)) %>%
   filter(vote == "Leave the EU") %>%
      ggplot(aes(x = endtime,
                        y= count,
                        group = country,
                        color = country)) +
      geom_line() + geom_point()
```



